---
title: "More about CAVAanalytics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=F, warning=FALSE}
library(CAVAanalytics)
```


The main purpose of CAVAanalytics is making the processing and visualization of multiple climate models simple. Additionally, CAVAanalytics offers an access point to download CORDEX-CORE climate models, the high resolution (25 Km) dinamically-downscaled climate models used in the IPCC report AR5. CAVAanalytics also offers access to state-of-the-art reanalyses datasets, such as W5E5 and ERA5. 

CAVAanalytics is built on top of climate4R and several tidyverse packages. The current available data that can be downloaded using the package are:

* CORDEX-CORE simulations (3 GCMs donwscaled with 2 RCMs for two RCPs)
* W5E5 and ERA5 reanalyses datasets

Available variables are:

* Daily maximum temperature (tasmax)
* Daily minimum temperature (tasmin)
* Daily precipitation (pr)
* Daily relative humidity (hurs)
* Daily wind speed (sfcWind)
* Daily solar radiation (rsds)

## Loading CORDEX-CORE data
Below we are automatically downloading all CORDEX-CORE simulations alongside the observational dataset W5E5 for Kenya. As you can see, it is extremely simple. We will see how to work with locally available models in another vignette.  
**Note that this process can take a while, depending on the size of the country and the number of years specified**

```{r, message=FALSE, warning=F}
# to download a specific area, set country=NULL and specify xlim and ylim arguments. 
kenya_exmp <- load_data(path.to.data = "CORDEX-CORE", country = "Kenya", variable = "tasmax", 
years.proj = 2030:2055, years.hist = 1980:2005, domain = "AFR-22", path.to.obs = "ERA5")

```

You can explore the output using attributes(kenya_exmp).

```{r,  message=F, warning=F}
# the main output of the loading function is a tibble with list column containing all the downloaded data
class(kenya_exmp[[1]])
colnames(kenya_exmp[[1]])
kenya_exmp[[1]]$experiment

# the column models_mbrs stands for model members. The structure of this element comes from the loadeR package. The data is contained in the data element, and as you can see, it has 6 members. CAVAanalytics automatically binds together multiple models and checks for temporal consistency.
dim(kenya_exmp[[1]]$models_mbrs[[1]]$Data)

```

## Observations

We can examine historical data using the loaded observational dataset, which, in this case, is ERA5. In this case, we are looking at the mean historical maximum temperature for the whole year. 

```{r, fig.height=4, fig.width=6,  message=F, warning=F}

kenya_obs <- observations(kenya_exmp, season=1:12)

plotting(kenya_obs, plot_titles = "Mean tasmax")

```

How many days above 40 °C occurred on average?

```{r, fig.height=4, fig.width=6,  message=F, warning=F}

kenya_obs.40 <- observations(kenya_exmp, season=1:12, uppert=40)

plotting(kenya_obs.40, plot_titles = "N.days > 40")

```


## Projections

What if you want to have a look at what will be the projected maximum temperature in Kenya for the 2030-2055 time period?
This is easy to do with CAVAanalytics:

```{r, fig.height=4, fig.width=6,  message=F, warning=F}
# you can specify the season. In this case we are looking at the whole year
proj_kenya <- projections(kenya_exmp, season=1:12)

# stat can take either mean or sd. It is used in combination with ensemble=T. Default is always mean
plotting(proj_kenya, ensemble=T, stat="mean", plot_titles = "mean tasmax")
```


```{r, fig.height=6, fig.width=10,  message=F, warning=F}

plotting(proj_kenya, ensemble=F, stat="mean", plot_titles = "mean tasmax")

```

In certain cases, focusing solely on mean temperatures might not provide the insights we need. Instead, we might be interested in understanding the frequency of specific conditions projected to occur. For instance, we may want to determine the number of days with maximum temperatures above 35 °C.


```{r, fig.height=4, fig.width=6,  message=F, warning=F}

proj_kenya_35 <- projections(kenya_exmp, season=1:12, uppert=35)

plotting(proj_kenya_35, ensemble=T, plot_titles = "N.days > 35", palette=c("white", "orange", "red", "darkred"))

```


When dealing with model biases, CAVAanalytics offers a straightforward solution by automatically implementing bias correction using the empirical quantile mapping method. This process allows us to correct for discrepancies between observed data and model outputs.

**it is essential to consider that bias correction with the W5E5 dataset will result in a change in spatial resolution to 50 Km. To maintain the original CORDEX-CORE resolution, it is crucial to perform bias correction using ERA5 instead**


```{r, fig.height=4, fig.width=6,  message=F, warning=F}
# bias correction is applied with the widely used empirical quantile method
proj_kenya_35bc <- projections(kenya_exmp, season=1:12, bias.correction = T, uppert=35)

plotting(proj_kenya_35bc, ensemble=T,  plot_titles = "N.days > 35 after bc", palette=c("white", "orange", "red", "darkred"), legend_range = c(0, 320))

```


## Climate change signal

While analyzing future periods can be beneficial, there are instances where we prioritize understanding delta changes, which represent the relative increase or decrease in mean maximum temperatures compared to a baseline. CAVAanalytics facilitates this analysis as well.

```{r, fig.height=6, fig.width=10,  message=F, warning=F}

ccs_kenya <- climate_change_signal(kenya_exmp, season=1:12)

plotting(ccs_kenya, ensemble=F, plot_titles = "Mean tasmax change", legend_range = c(-3, 3), palette = c("blue", "cyan", "green", "white", "yellow" ,"orange", "red"))

```

## Trends
While examining delta changes and absolute values for projections provides valuable information, it may not offer insights into temporal trends. However, using the 'trends' function in CAVAanalytics allows us to address this limitation. For instance, we can determine if Kenya has experienced or is projected to experience a noteworthy rise in yearly maximum temperatures.

```{r,  fig.height=4, fig.width=6,  message=F, warning=F}
# the observation argument allows us to study trends for the future or for the past. Note that for observation = T, we need to have specified path.to.obs because this option does not work with the historical experiment of climate models but with observation

trd_kenya.obs <- trends(kenya_exmp, season=1:12,observation = T)

plotting(trd_kenya.obs, ensemble=F, plot_titles = "°C/year", frequencies = F, spatial_aggr = F)

```

We can look at trend for the future period as well by changing the observation argument to FALSE

```{r,  fig.height=6, fig.width=10,  message=F, warning=F}

trd_kenya.prj <- trends(kenya_exmp, season=1:12,observation = F)

plotting(trd_kenya.prj, ensemble=F, plot_titles = "°C/year", frequencies = F, spatial_aggr = F)

```

Another feature of the trends function is that you can look at the global significance of the ensemble. The plot below shows you the number of models that agrees in a certain change (increase or decrease) while the global p.value is calculated based on the [manylm function of mvabund](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/j.2041-210X.2012.00190.x). 

```{r,  fig.height=4, fig.width=6,  message=F, warning=F}

plotting(trd_kenya.prj, ensemble=T, plot_titles = "N. members direction change", frequencies = F, spatial_aggr = F, palette = c("blue", "green", "white", "yellow", "red"))

```


While we can visualize the results spatially, we can also look at temporal trends. 

```{r, fig.width=8,  message=F, warning=F}

plotting(trd_kenya.prj, ensemble=T, plot_titles = "°C/year", spatial_aggr = T, frequencies = F)

```

In addition to performing spatial averages and visualizing trends, we can gain valuable insights by examining the distribution of annual mean tasmax (maximum temperature) across Kenya using the specified frequencies, denoted as 'TRUE.' This approach is particularly valuable because many analyses tend to aggregate data either spatially or temporally, potentially obscuring certain patterns. By utilizing the 'frequencies=T' option, we ensure a comprehensive understanding of both spatial and temporal attributes, allowing us to uncover hidden patterns in the data."

```{r, fig.width=10,  message=F, warning=F}

plotting(trd_kenya.prj, ensemble=T, spatial_aggr = F, frequencies = T, n.groups = 2)

```


