---
title: "The power of patchwork with CAVAanalytics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=F, warning=FALSE}
library(CAVAanalytics)
library(patchwork)
library(ggplot2)
library(magrittr)
```

In the [getting started article](https://risk-team.github.io/CAVAanalytics/articles/Introduction.html), we have already explored some of the fundamental features of CAVAanalytics. Now, let's take our understanding to the next level as we delve into the full potential of CAVAanalytics, especially when synergized with the powerful library [patchwork](https://patchwork.data-imaginist.com/) for crafting stunning plots.
So, without further ado, let's dive right in by loading the data.


```{r, message=F, warning=F}
# to download a specific area, set country=NULL and specify xlim and ylim arguments. 
kenya_exmp <- load_data(path.to.data = "CORDEX-CORE", country = "Kenya", variable = "tasmax", 
years.proj = 2030:2060, years.hist = NULL, domain = "AFR-22")

```

## Composite seasonal plot

```{r,  warning=F, message=FALSE, fig.height=8, fig.width=10}
# Here we use map from purrr to create seasonal plots for projections. Remember that the output of plotting is a ggplot object, hence you can add anything to it or change the scale, theme, etc

sns_proj <- purrr::map(list(1:3, 4:6, 7:9, 10:12), ~ projections(kenya_exmp, season = .x) %>%
  plotting(., plot_titles = "mean tasmax °C", ensemble = T, legend_range = c(20, 40))+
    ggtitle(paste0("Season ", glue::glue_collapse(.x, "-")))+ # add title
    theme(plot.title = element_text(hjust = 0.5))) # center title

wrap_plots(sns_proj)


```

These plots have the same legends, thus we might want to collect them. 

```{r,  warning=F, message=FALSE, fig.height=8, fig.width=10}

(composite <- wrap_plots(sns_proj)+plot_layout(guides = "collect") & theme(legend.position = 'bottom'))

```

Now we add the the previous seasonal plot the overall trends

```{r,  warning=F, message=FALSE, fig.height=9, fig.width=8}

ind.models <- trends(kenya_exmp, season = 1:12, observation = F) %>%
  plotting(., plot_titles = "°C/year", spatial_aggr = T, frequencies = F, ensemble=T, palette = c("blue", "red"))+
  ggtitle("Season 1-12")+ # add title
  theme(plot.title = element_text(hjust = 0.5))  # center title

composite/ind.models

```
In a similar manner, other composite plots can be made, for example to combine different time frames instead of different seasons. 

